---
title: "810project"
author: "Chiebuka Onwuzurike (U15815811)"
date: "2/21/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r}
library(data.table) 
library(ggplot2) 
library(ggthemes) 
library(glmnet) 
theme_set(theme_bw())
```
```{r}
dd <- fread("/Users/chieb/Documents/Coding/School_Work/BA 810/Melbourne_housing_FULL.csv")
# data size and structure
# dim(dd)
# str(dd)
```
## Data Cleaning
```{r}
# view missing values
sum(is.na(dd))
colSums(is.na(dd))
```
```{r}
# removing missing values
dd <- dd[!(is.na(dd$Price))]
dd <- dd[,BuildingArea:=NULL]
dd <- dd[!is.na(dd$Landsize)]
dd <- dd[!is.na(dd$YearBuilt)]
dd$Car[which(is.na(dd$Car))] <- 0
colSums(is.na(dd))
dim(dd)
```
## Features Engineering
```{r}
dd$Type[dd$Type == 'h'] <- 1
dd$Type[dd$Type == 't'] <- 2
dd$Type[dd$Type == 'u'] <- 3
dd$Type <- as.factor(dd$Type)
```
```{r}
dd$Method[dd$Method == 'PI'] <- 1
dd$Method[dd$Method == 'PN'] <- 2
dd$Method[dd$Method == 'S'] <- 3
dd$Method[dd$Method == 'SA'] <- 4
dd$Method[dd$Method == 'SN'] <- 5
dd$Method[dd$Method == 'SP'] <- 6
dd$Method[dd$Method == 'SS'] <- 7
dd$Method[dd$Method == 'VB'] <- 8
dd$Method[dd$Method == 'W'] <- 9
dd$Method <- as.factor(dd$Method)
```
```{r}
dd$Regionname[dd$Regionname == 'Eastern Metropolitan'] <- 1
dd$Regionname[dd$Regionname == 'Eastern Victoria'] <- 2
dd$Regionname[dd$Regionname == 'Northern Metropolitan'] <- 3
dd$Regionname[dd$Regionname == 'Northern Victoria'] <- 4
dd$Regionname[dd$Regionname == 'South-Eastern Metropolitan'] <- 5
dd$Regionname[dd$Regionname == 'Southern Metropolitan'] <- 6
dd$Regionname[dd$Regionname == 'Western Metropolitan'] <- 7
dd$Regionname[dd$Regionname == 'Western Victoria'] <- 8
dd$Regionname <- as.factor(dd$Regionname)
```
```{r}
dd$Date <- as.Date(dd$Date,"%d/%m/%Y")
```

## Splitting data into training and test set by creating new column to identify test. Split 80/20
```{r}
dd[, test:=0]
dd[sample(nrow(dd), round(nrow(dd)*.2, digits = 0)), test:=1]
dd.test <- dd[test==1]
dd.train <- dd[test==0]
```

colnames(dd)

## Formulas
```{r}
f1 <- as.formula(Price ~ Rooms + Type + Method + Distance + Bedroom2 + Bathroom + Car + Regionname)
# Unused Variables: SellerG + Subrub +  Date  + Postcode + CouncilArea
```

## Creating Variables

```{r} 
x1.train <- model.matrix(f1, dd.train)[, -1]
x1.test <- model.matrix(f1, dd.test)[, -1]
y.train <- dd.train$Price
y.test <- dd.test$Price
```

dim(x1.train)
length(y.train)
dim(x1.test)
length(y.test)


length(unique(dd$Postcode))

length(unique(dd$CouncilArea))

 ## Baseline Linear Regression
```{r}
fit.lm <- lm(f1, dd.train)

yhat.train.lm <- predict(fit.lm)
mse.train.lm <- mean((y.train - yhat.train.lm)^2)
mse.train.lm

yhat.test.lm <- predict(fit.lm, dd.test)
mse.test.lm <- mean((y.test - yhat.test.lm)^2)
mse.test.lm
```

## Ridge Regression with 10 fold cross validation
```{r}
fit.ridge <- cv.glmnet(x1.train, y.train, alpha = 0, nfolds = 10)

yhat.train.ridge <- predict(fit.ridge, x1.train, s = fit.ridge$lambda.min)
mse.train.ridge <- mean((y.train - yhat.train.ridge)^2)
mse.train.ridge


yhat.test.ridge <- predict(fit.ridge, x1.test, s = fit.ridge$lambda.min)
mse.test.ridge <- mean((y.test - yhat.test.ridge)^2)
mse.test.ridge
```

## Lasso Regression with 10 fold cross validation
```{r}
fit.lasso <- cv.glmnet(x1.train, y.train, alpha = 1, nfolds = 10)


yhat.train.lasso <- predict(fit.lasso, x1.train, s = fit.lasso$lambda.min)
mse.train.lasso <- mean((y.train - yhat.train.lasso)^2)
mse.train.lasso

yhat.test.lasso <- predict(fit.lasso, x1.test, s = fit.lasso$lambda.min)
mse.test.lasso <- mean((y.test - yhat.test.lasso)^2)
mse.test.lasso
```

plot_coeffs <- function(model) {
      coeffs <- coefficients(model)
      mp <- barplot(coeffs, col="#3F97D0", xaxt='n', main="Regression Coefficients")
      lablist <- names(coeffs)
      text(mp, par("usr")[3], labels = lablist, srt = 45, adj = c(1.1,1.1), xpd = TRUE, cex=0.6)
    }

plot_coeffs(x1.train)